### 2017-06-16 manuscript/manuscript.Rmd
initial commit

---
title: Symbiont shuffling linked to differential photochemical efficiency of *Symbiodinium* in three Caribbean reef corals
author: "Ross Cunning"
date: "6/7/2017"
output: html_document
---

Title ideas:

Symbiont shuffling linked to differential photochemical efficiency of *Symbiodinium* in three Caribbean reef corals