### 2017-06-16 supp/supp.Rmd
add figure legends, trim extra

---
title: "Symbiont shuffling linked to differential photochemical efficiency of *Symbiodinium* in three Caribbean reef corals"
output: html_document
---

```{r setup, include=FALSE}
# Load libraries
library(MASS)
library(effects)
library(reshape2)
library(scales)
library(lme4)
library(lmerTest)
library(lsmeans)
library(lattice)
library(dplyr)

# Set knit options
knitr::opts_chunk$set(warning=F, message=F, fig.height=5, fig.width=5)
knitr::opts_knit$set(root.dir = "..")

# Polygon drawing function
addpoly <- function(x, y1, y2, col=alpha("lightgrey", 0.8), ...){
  ii <- order(x)
  y1 <- y1[ii]
  y2 <- y2[ii]
  x <- x[ii]
  polygon(c(x, rev(x)), c(y1, rev(y2)), col=col, border=NA, ...)
}
```

# Import data
```{r load_data}
# Load all data from file
alldata <- read.csv("data/master.csv")
alldata <- alldata[-360, ]  # Remove duplicate Mc9_2 entry with incorrect Fv/Fm recov2 value
alldata$Species <- factor(alldata$Species, levels=c("Of", "Ss", "Mc"))

# Get data frame with all Ofav and Ssid data
SsOf <- subset(alldata, Species!="Mc")

# Get data frame with Mcav repeat bleaching data (=second bleaching)
Mc <- subset(alldata, BleachLvl2=="Heat" & Species=="Mc")
## Re-level Mcav data for merging with Ofav and Ssid data
## Set data after recovery from first bleaching to the "initial" factor levels
Mc$PropD.init <- Mc$PropD.recov  
Mc$MaxY.init <- Mc$MaxY.recov
## Set data after the second bleaching to the "bleached" factor levels
Mc$BleachLvl <- as.character(Mc$BleachLvl2)
Mc$PropD.bleach <- Mc$PropD.bleach2
Mc$MaxY.bleach <- Mc$MaxY.bleach2
## Set BleachLvl to "Med" to match Ofav and Ssid factor level
Mc$BleachLvl[Mc$BleachLvl=="Heat"] <- "Med"
## Set data after second recovery to the "recove#d95f02" factor levels
Mc$PropD.recov <- Mc$PropD.recov2
Mc$MaxY.recov <- Mc$MaxY.recov2

# Merge data from all three coral species together
bleach3c <- rbind(SsOf, Mc)[, c("Species", "Colony", "Core", "Total.init", "PropD.init", "MaxY.init", "BleachLvl", "Total.bleach", "PropD.bleach", "MaxY.bleach")]
bleach3c <- droplevels(bleach3c)

# Melt data frame so that "time" (i.e., initial vs. bleached) can be used as factor
bleachmelt <- melt(bleach3c, id.vars = c("Species", "Colony", "Core", "BleachLvl"))
bleachmelt <- cbind(bleachmelt, 
                    colsplit(bleachmelt$variable, pattern="\\.", 
                             names=c("resp", "time")))
bleachdat <- dcast(bleachmelt, 
                   formula = Species + Colony + Core + BleachLvl + time ~ resp, value.var="value")
bleachdat$time <- factor(bleachdat$time)
```

# Analyze photochemical efficiency (Fv/Fm)
```{r analyze_fvfm}
# Test of slopes of bleached FvFm ~ PropD are different for different BleachLvls
testmod <- lmer(MaxY ~ PropD * Species * BleachLvl + (1|Colony), 
                data=subset(bleachdat, time=="bleach" & Species!="Mc"))
anova(testmod)
plot(Effect(c("PropD", "BleachLvl", "Species"), testmod), multiline=T, z.var="BleachLvl")
slopes <- lstrends(testmod, specs=c("BleachLvl", "Species"), var="PropD")
cld(contrast(slopes, by="Species"))
## No difference in photochemical advantage of D under stress among different BleachLvl's. Therefore, can omit BleachLvl from statistical model for calculating photochemical advantage of D

# Fit model for Fv/Fm
bleachmod <- lmer(MaxY ~ PropD * time * Species + (1|Colony), bleachdat)
anova(bleachmod)
select <- lmerTest::step(bleachmod)$model
bleachmod <- lmer(formula(select), data=bleachdat)
anova(bleachmod)

plot(Effect(c("PropD", "time", "Species"), bleachmod), multiline=T, ci.style="bands")

# Compare slopes of Fv/Fm vs. PropD among species by time
lst <- lstrends(bleachmod, specs=c("Species", "time"), var="PropD")
cld(contrast(lst, by="time"))
plot(lst, horizontal=F)

# Make data frame to collect fitted values and confidence intervals
pred <- expand.grid(Species=c("Of", "Ss", "Mc"), 
                    PropD=seq(0,1,len=101), 
                    time=c("init", "bleach"))
# Get fitted Fv/Fm values from model
pred$fit <- predict(bleachmod, pred, re.form=NA)

# Bootstrap model to get confidence intervals on fitted values
bootfit <- bootMer(bleachmod, FUN=function(x) predict(x, pred, re.form=NA), nsim=999)
# Extract 90% confidence interval on predicted values
pred$lci <- apply(bootfit$t, 2, quantile, 0.05)
pred$uci <- apply(bootfit$t, 2, quantile, 0.95)
```

## Plot Figure 1
```{r Fig1}
png("figures/Fig1.png", width=4, height=2.5, units="in", res=300)
par(mfrow=c(1, 3), mar=c(3,3,1.1,0), mgp=c(1.5, 0.1, 0), 
    tcl=0.2, cex.axis=0.8, xaxs="i")
plot(NA, xlim=c(0, 1), ylim=c(0.2, 0.55),
     xlab="Prop. clade D", ylab="Fv/Fm")
mtext("A.", adj=-0.2, font=2)
with(subset(pred, Species=="Of" & time=="init"), {
  lines(fit ~ PropD, lty=1, lwd=1)
  addpoly(PropD, lci, uci, col=alpha("#1b9e77", 0.5))
})
with(subset(pred, Species=="Of" & time=="bleach"), {
  lines(fit ~ PropD, lty=2, lwd=2)
  addpoly(PropD, lci, uci, col=alpha("#1b9e77", 0.5))
})

par(mar=c(3,1.5,1.1,1.5))
plot(NA, xlim=c(0, 1), ylim=c(0.2, 0.55),
     xlab="Prop. clade D", ylab="")
mtext("B.", adj=-0.2, font=2)
with(subset(pred, Species=="Ss" & time=="init"), {
  lines(fit ~ PropD, lty=1, lwd=1)
  addpoly(PropD, lci, uci, col=alpha("#7570b3", 0.5))
})
with(subset(pred, Species=="Ss" & time=="bleach"), {
  lines(fit ~ PropD, lty=2, lwd=2)
  addpoly(PropD, lci, uci, col=alpha("#7570b3", 0.5))
})

par(mar=c(3,0,1.1,3))
plot(NA, xlim=c(0, 1), ylim=c(0.2, 0.55),
     xlab="Prop. clade D", ylab="")
mtext("C.", adj=-0.2, font=2)
with(subset(pred, Species=="Mc" & time=="init"), {
  lines(fit ~ PropD, lty=1, lwd=1)
  addpoly(PropD, lci, uci, col=alpha("#d95f02", 0.5))
})
with(subset(pred, Species=="Mc" & time=="bleach"), {
  lines(fit ~ PropD, lty=2, lwd=2)
  addpoly(PropD, lci, uci, col=alpha("#d95f02", 0.5))
})
dev.off()
```
![Figure 1. Photochemical efficiency as a function of the proportion of clade D *Symbiodinium*. Solid lines represent ambient conditions, and dotted lines represent the relationship under heat stress (32°C). Shaded regions represent 90% confidence intervals around fitted values.](../figures/Fig1.png)

```{r Fig2}
# Plot Advantage of D vs. other symbiont in each coral under
# normal conditions and under heat stress (in Fv/Fm units)

# Get ADH and ADA values (Advantage of D under Heat and under Ambient conditions)
AD <- summary(lst)
ADm <- melt(AD)
ADm$time <- factor(ADm$time, levels=c("init", "bleach"))
ADf <- as.matrix(dcast(subset(ADm, variable=="PropD.trend"), time ~ Species)[,-1])
ADu <- as.matrix(dcast(subset(ADm, variable=="upper.CL"), time ~ Species)[, -1])
ADl <- as.matrix(dcast(subset(ADm, variable=="lower.CL"), time ~ Species)[, -1])

png("figures/Fig2.png", width=3, height=3, units="in", res=300)
par(mar=c(1.8,3.2,1,0.8), mgp=c(1.5, 0.25, 0), tcl=-0.2, cex.axis=0.8)
bars <- barplot(ADf, beside = T, ylim=c(-0.2,0.2),
                names.arg=c("Ofav", "Ssid", "Mcav"),
                col=c("#1b9e77", alpha("#1b9e77", 0.3),
                      "#7570b3", alpha("#7570b3", 0.3),
                      "#d95f02", alpha("#d95f02", 0.3)))
arrows(matrix(bars), matrix(ADl), matrix(bars), matrix(ADu), 
       length = 0.05, lwd=1, xpd=NA, code=3, angle=90)
mtext(side=2, "Photochemical advantage of clade D", 
      line=1.5, cex=0.8)
#legend("topleft", pch=22, pt.bg=c("gray80", "gray20"), y.intersp=1.2,
#       pt.cex=2, cex=0.6, inset=c(0.02, -0.05), xpd=NA, bty = "n", 
#       legend = c(expression("Heat stress"~"(A"["Heat"]^"D"~")"),
#                  expression("Ambient"~"(A"["Amb"]^"D"~")")))
abline(h=0)
dev.off()
```
![Figure 2. Photochemical advantage of clade D symbionts in each coral species under ambient conditions (darker bars) and under heat stress (lighter bars). Values indicate the difference in Fv/Fm of corals with all and no clade D symbionts (i.e., slopes of the lines in Fig. 1), with error bars indicating 95% confidence intervals.](../figures/Fig2.png)

# Symbiont shuffling after bleaching
```{r analyze_symshuff}
recovdat <- rbind(SsOf, Mc)[, c("Species", "Colony", "Core", "Total.init", "PropD.init", "MaxY.init", "BleachLvl", "RecovTemp", "Total.recov", "PropD.recov", "MaxY.recov")]

propDmod <- glm(PropD.recov ~ PropD.init + Species*BleachLvl*RecovTemp,
                 data=recovdat, family=quasibinomial)
summary(propDmod)  # Model summary
dropterm(propDmod, test="F")  # Partial F-tests

# Get fitted values for PropD.recov, averaging across RecovTemp levels
eff <- effect(c("PropD.init", "Species", "BleachLvl"), propDmod, 
              xlevels=list(PropD.init=seq(0, 1, 0.01)))

# Get all fitted values and subsets for each bleaching level
res <- droplevels(data.frame(eff))
res$Species <- factor(res$Species, levels=c("Of", "Ss", "Mc"))
res.Med <- subset(data.frame(eff), BleachLvl=="Med")
res.Low <- subset(data.frame(eff), BleachLvl=="Low")
res.High <- subset(data.frame(eff), BleachLvl=="High")

# Get AUC for fitted values, lower and upper confidence limits
auc <- aggregate(res[, c("fit", "lower", "upper")], 
                 by=list(Species=res$Species, BleachLvl=res$BleachLvl), 
                 FUN=function(x) (mean(x)-0.5)/0.5)
auc.list <- split(auc, list(auc$BleachLvl))
```

```{r Fig3}
# Figure 3
png("figures/Fig3.png", width=7, height=3.5, units="in", res=300)
par(mfrow=c(1, 2), mar=c(3,3,1,1), mgp=c(1.5, 0.1, 0), tcl=0.2, cex.axis=0.8)
## Figure 3A: Final vs. initial proportion clade D
plot(NA, xlim=c(0,1), ylim=c(0,1), xaxs="i", yaxs="i",
     xlab="Prop. D before bleaching",
     ylab="Prop. D after recovery")
# Plot lines for low bleaching
with(subset(res.Low, Species=="Mc"), {
  lines(fit ~ PropD.init, lwd=2, col="#d95f02", xpd=NA)
  addpoly(PropD.init, lower, upper, col = alpha("#d95f02", 0.4))
  Mc.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
with(subset(res.Low, Species=="Of"), {
  lines(fit ~ PropD.init, lwd=2, col="#1b9e77", xpd=NA)
  addpoly(PropD.init, lower, upper, col = alpha("#1b9e77", 0.4))
  Of.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
with(subset(res.Low, Species=="Ss"), {
  lines(fit ~ PropD.init, lwd=2, col="#7570b3", xpd=NA)
  addpoly(PropD.init, lower, upper, col = alpha("#7570b3", 0.4))
  Ss.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
abline(0,1, lty=1, lwd=0.5)
# Plot lines for Med bleaching
with(subset(res.Med, Species=="Mc"), {
  lines(fit ~ PropD.init, lwd=2, col="#d95f02", xpd=NA, lty=2)
  addpoly(PropD.init, lower, upper, col = alpha("#d95f02", 0.4))
  Mc.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
with(subset(res.Med, Species=="Of"), {
  lines(fit ~ PropD.init, lwd=2, col="#1b9e77", xpd=NA, lty=2)
  addpoly(PropD.init, lower, upper, col = alpha("#1b9e77", 0.4))
  Of.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
with(subset(res.Med, Species=="Ss"), {
  lines(fit ~ PropD.init, lwd=2, col="#7570b3", xpd=NA, lty=2)
  addpoly(PropD.init, lower, upper, col = alpha("#7570b3", 0.4))
  Ss.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
# Plot lines for High bleaching
with(subset(res.High, Species=="Mc"), {
  lines(fit ~ PropD.init, lwd=2, col="#d95f02", xpd=NA, lty=3)
  addpoly(PropD.init, lower, upper, col = alpha("#d95f02", 0.4))
  Mc.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
with(subset(res.High, Species=="Of"), {
  lines(fit ~ PropD.init, lwd=2, col="#1b9e77", xpd=NA, lty=3)
  addpoly(PropD.init, lower, upper, col = alpha("#1b9e77", 0.4))
  Of.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
with(subset(res.High, Species=="Ss"), {
  lines(fit ~ PropD.init, lwd=2, col="#7570b3", xpd=NA, lty=3)
  addpoly(PropD.init, lower, upper, col = alpha("#7570b3", 0.4))
  Ss.auc <<- c((mean(fit)-0.5)/0.5, (mean(lower)-0.5)/0.5, (mean(upper)-0.5)/0.5)
})
legend("bottomright", bty="n", lty=c(1,2,3), lwd=2, inset=c(0.05, 0),
       legend=c("7 days", "10 days", "14 days"), cex=0.7)
legend("bottomright", bty="n", pch=15, inset=c(0.04, 0.175),
       legend=c("Mcav", "Ssid", "Ofav"), cex=0.7, pt.cex=1.5, x.intersp=1.5,
       col=c("#d95f02", "#7570b3", "#1b9e77"))
#text(par("usr")[1]-par("mai")[2]/2, par("usr")[4], "A.", font=2, xpd=NA)
#text(-0.2, par("usr")[4], "A.", font=2, xpd=NA)
mtext("A.", adj=-0.2, font=2)

# Figure 3B: AUC for each species at each BleachLvl
plot(NA, ylim=c(-1,1), xlim=c(0.5,3.5), xaxs="i", yaxs="i", xaxt="n", 
     xlab="Species", ylab="Symbiont shuffling")
axis(side=1, at=1:3, labels=c("Of", "Ss", "Mc"))
abline(h=0, lty=1, lwd=0.5)
with(auc.list$Low, {
  arrows(0.8:2.8, lower, 0.8:2.8, upper, code=3, length=0.075, angle=90, xpd=NA)
  points(0.8:2.8, fit, pch=15, col=c("#1b9e77", "#7570b3", "#d95f02"), xpd=NA)
})
with(auc.list$Med, {
  arrows(1:3, lower, 1:3, upper, code=3, length=0.075, angle=90, xpd=NA)
  points(1:3, fit, pch=17, col=c("#1b9e77", "#7570b3", "#d95f02"), xpd=NA)
})
with(auc.list$High, {
  arrows(1.2:3.2, lower, 1.2:3.2, upper, code=3, length=0.075, angle=90, xpd=NA)
  points(1.2:3.2, fit, pch=16,  col=c("#1b9e77", "#7570b3", "#d95f02"), xpd=NA)
})
legend("bottomright", bty="n", inset=c(0.05, 0.1),
       pch=c(15, 17, 16), col="black", cex=0.7,
       legend=c("7 days", "10 days", "14 days"))

arrows(par("usr")[2]*0.95, c(0.02, -0.02), 
       par("usr")[2]*0.95, c(0.2, -0.2), length=0.1)
text(par("usr")[2]*0.95, c(0.08, -0.1), pos=2, cex=0.7,
     labels=c("more clade D", "less clade D"))
#text(par("usr")[1]-par("mai")[2]/2, par("usr")[4], "B.", font=2, xpd=NA)
mtext("B.", adj=-0.2, font=2)
dev.off()
```
![Figure 3. Symbiont shuffling in three coral species. Panel **(A)** shows the proportion of clade D *Symbiodinium* before bleaching and after recovery, with the diagonal line of identity representing no change in symbiont proportions. Fitted values for each species (indicated by color) and heat stress duration (indicated by line type) are plotted with shaded areas representing 95% confidence intervals. Panel **(B)** shows the integrated symbiont shuffling response (from before bleaching to after recovery) for each species and bleaching treatment, calculated as the area above or below the diagonal line of identity, normalized to a range of -1 (i.e., a complete loss of clade D symbionts) to +1 (i.e., complete dominance by clade D symbionts).](../figures/Fig3.png)

```{r Fig4}

# Model shuffling as a function of photochemical advantage and heat stress

# Get AD (photochemical Advantage of D, =slopes of Fv/Fm~propD lines)
AD <- summary(lst)[,1:3]
df <- merge(AD, auc)
## ADA = Advantage of D under Ambient conditions (not relative -- in Fv/Fm units)
df$ADA <- with(subset(AD, time=="init"), 
                      PropD.trend[match(df$Species, Species)])
df$ADH <- with(subset(AD, time=="bleach"), 
                      PropD.trend[match(df$Species, Species)])
# Change column name for fitted shuffling response to 'shuffle'
colnames(df) <- stringr::str_replace(colnames(df), "fit", "shuffle")
# Convert BleachLvl factor to # days of heat stress
df$days <- ifelse(df$BleachLvl=="Low", 7, ifelse(df$BleachLvl=="Med", 10, 14))
# Rescale shuffling response range from [-1,1] to [0,1]
df$shuffle.scaled <- scales::rescale(df$shuffle, from=c(-1, 1), to=c(0, 1))
# Create variables for products of Fv/Fm differences (AD metrics) and days of heat
df$ADH.days <- df$ADH * df$days
df$ADA.days <- df$ADA * df$days

# P#d95f02ict shuffling response based on relative Fv/Fm value and BleachLvl
## relative Fv/Fm at ambient temperatures ("normal") is perfectly correlated with relative Fv/Fm at stress temperatures ("stress") so only one variable can be used in model.  Or, should the slope of the Fv/Fm line be used, rather than the "relative difference"? (makes very little difference whether relative or slope is used.) *slope needs to be used because relative is different for diff BleachLvl levels, but slope is the same!

png("figures/Fig4.png", width=7, height=3.5, units="in", res=300)
par(mfrow=c(1,2), mgp=c(1.5,0.1,0), mar=c(3,3,1,1), tcl=0.2, cex.axis=0.8)
# Orthogonal model
glmod <- glm(shuffle.scaled ~ ADH + days, data=df, family="quasibinomial")
anova(glmod, test="F")

# Get fitted values and plot shuffling ~ days for different relative advantages of D
stresslevs <- c(0.033, 0.05, 0.07, 0.103, 0.163)
cols <- c("#1b9e77", "gray40", "gray40", "#7570b3", "#d95f02")
resdf <- expand.grid(days=7:14, ADH=stresslevs)
resdf$fit <- predict(glmod, newdata=resdf, type="response")
resdf$fit.bt <- rescale(resdf$fit, from=c(0,1), to=c(-1,1))
plot(NA, xlim=c(7,14), ylim=c(-1,1), yaxs="r",
     xlab="Days of heat stress",
     ylab="Symbiont shuffling")
mtext("A.", adj=-0.2, font=2)
reslist <- split(resdf, f=resdf$ADH)
for (i in 1:length(reslist)) {
  with(reslist[[i]], lines(fit.bt ~ days, col=cols[i], lwd=2, xpd=NA))
}
abline(h=0, lwd=0.5)
# Get positions for text annotations
angles <- unlist(lapply(reslist, function(x) with(x, fit.bt[2]-fit.bt[1])))
heights <- unlist(lapply(reslist, function(x) with(x, fit.bt[1])))
ADHeat <- expression("A"["Heat"]^"D"~"=")
for (i in 1:length(stresslevs)) {
  text(labels=bquote(.(parse(text=paste(ADHeat,"~",stresslevs[i],sep=""))) ),
       x=7, y=heights[i]-0.06, adj=0, srt=angles[i]*200, cex=0.6)
}
legend("bottomright", bty="n", inset=c(0.05, 0.1),
       lty=1, lwd=2, col=c("#d95f02", "#7570b3", "#1b9e77"), cex=0.7,
       legend=c("Mcav", "Ssid", "Ofav"))

## Panel B
glmod2 <- glm(shuffle.scaled ~ ADH.days, family="quasibinomial", data=df)
predicted <- predict(glmod2, 
                     newdata=data.frame(ADH.days=seq(0,3,0.1)), 
                     type="response")
with(summary(glmod2), 1-(deviance(glmod2)/null.deviance))

plot(df$shuffle ~ df$ADH.days, ylim=c(-1,1), yaxs="r", xpd=NA,
     pch=c(16,15,17)[as.numeric(as.factor(df$BleachLvl))],
     xlab=expression("A"["Heat"]^"D"~"x (Days of heat stress)"),
     ylab="", mgp=c(1.8,0.1,0),
     col=c("#d95f02", "#1b9e77", "#7570b3")[as.numeric(factor(df$Species))])
mtext("Symbiont shuffling", side=2, line=1.5)
mtext("B.", adj=-0.2, font=2)
lines(seq(0,3,0.1), rescale(predicted, from=c(0,1), to=c(-1,1)), lwd=1)
abline(h=0, lwd=0.5)
text(par("usr")[2]*0.95, c(0.08, -0.1), pos=2, cex=0.7,
     labels=c("more clade D", "less clade D"))
arrows(par("usr")[2]*0.95, c(0.02, -0.02), 
       par("usr")[2]*0.95, c(0.2, -0.2), length=0.1)
legend("bottomright", bty="n", inset=c(0.05, 0.1),
       pch=c(15, 17, 16), col="black", cex=0.7,
       legend=c("7 days", "10 days", "14 days"))
dev.off()
```
![Figure 4. Symbiont shuffling as a function of the photochemical advantage of clade D and heat stress duration. Panel **(A)** shows the symbiont shuffling response (as in Fig. 3B) as a function of heat stress duration (days) and the photochemical advantage of clade D under heat stress (A^D^~Heat~), with lines corresponding to the A^D^~Heat~ values for the three coral species studied (indicated by color), and two hypothetical intermediate values (gray lines). Panel **(B)** shows that the symbiont shuffling responses observed across all three species are predicted well (psuedo-R^2^=0.93) by the time-integrated photochemical advantage of clade D under heat stress.](../figures/Fig4.png)