### 2017-07-14 manuscript/manuscript.Rmd
AB edits

---
title: Symbiont shuffling linked to differential photochemical dynamics of *Symbiodinium*
  in three Caribbean reef corals
author:
- name: Ross Cunning
  affiliation: MBE
  footnote: Corresponding Author
  email: ross.cunning@gmail.com
- name: Rachel N. Silverstein
  affiliation: MBE,MWK
- name: Andrew C. Baker
  affiliation: MBE

bibliography: library.bib
output:
  rticles::elsevier_article:
    fig_caption: yes
  word_document: null
csl: coral-reefs.csl
address:
- address: Department of Marine Biology and Ecology, Rosenstiel School of Marine and
    Atmospheric Science, University of Miami, 4600 Rickenbacker Causeway, Miami, FL
    33149, USA
  code: MBE
- address: Miami Waterkeeper, 12568 N Kendall Drive, Miami, FL 33185, USA
  code: MWK
---
```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(lmerTest)
library(MASS)
```

```{r load_analysis}
load("analysis/analysis.RData")

ncolsperspp <- with(bleach3c, aggregate(Colony ~ Species, 
                                        FUN=function(x) length(unique(x))))
nOfavcol <- ncolsperspp[1, 2]
nSsidcol <- ncolsperspp[2, 2]
nMcavcol <- ncolsperspp[3, 2]

ncorepercol <- with(bleach3c, table(Species, Colony))
ncorepercol[ncorepercol==0] <- NA
rangeOfav <- range(ncorepercol[1, ], na.rm=T)
rangeSsid <- range(ncorepercol[2, ], na.rm=T)
rangeMcav <- range(ncorepercol[3, ], na.rm=T)

ncoreperspp <- with(bleach3c, table(Species))
nOfavcore <- ncoreperspp["Of"]
nSsidcore <- ncoreperspp["Ss"]
nMcavcore <- ncoreperspp["Mc"]

```

# Abstract
  Dynamic symbioses with functionally diverse dinoflagellate algae in the genus *Symbiodinium* may allow some reef corals to alter their phenotypes through 'symbiont shuffling', or changes in symbiont community composition. In particular, corals have been shown to become more bleaching-resistant by increasing the relative abundance of thermally tolerant *Symbiodinium* in clade D after bleaching. Despite the immediate relevance of this phenomenon to corals living in warming oceans -- and to interventions aimed at boosting coral resilience -- the mechanisms governing how, why, and when symbiont shuffling occurs are still poorly understood. Here, we performed controlled thermal bleaching and recovery experiments on three species of Caribbean reef-building corals, and show that the degree of symbiont shuffling is determined by (1) the difference in photochemical efficiency (*F~v~/F~m~*) of co-occurring symbionts under stress, and (2) the duration of stress exposure. The photochemical advantage of D1a (*S. trenchii*) under stress was greatest in *Montastraea cavernosa*, intermediate in *Siderastrea siderea*, and lowest in *Orbicella faveolata*, and correlated positively with the magnitude of increase in D1a. In holobionts where D1a had less of an advantage (i.e., a smaller difference in *F~v~/F~m~*), longer periods of heat stress were required to elicit commensurate increases in D1a. These findings help elucidate general rules governing the process of symbiont shuffling in reef corals, which will help refine predictions of coral responses to environmental change, and inform interventions to manipulate symbiont communities to enhance coral resilience.

# Introduction

Coral reef ecosystems are declining globally at an alarming rate, owing primarily to mass coral bleaching caused by anthropogenic increases in sea surface temperature [@Hughes:2017bs]. Heat-induced bleaching, or the loss of corals' nutritional symbionts (*Symbiodinium* spp.; @Jokiel:1977p7353), often leads to coral mortality, which precipitates reef ecosystem degradation [@Baker:2008p3421].

However, depending on the severity of bleaching, some corals may recover their *Symbiodinium* communities within weeks to months [@Jokiel:1977p7353; @Jones:1997p3499; @Cunning:2016jb], and ultimately return to a healthy state. Moreover, the loss and subsequent re-establishment of the symbiotic algal community may provide an opportunity for change in its composition, i.e., the relative abundance of different *Symbiodinium* types [@Buddemeier:1993p3; @Baker:2001p42]. When such 'symbiont shuffling' [@Baker:2003p200] occurs, the new configuration of the *Symbiodinium* assemblage may be better suited to tolerate future bleaching stressors [@Berkelmans:2006p191; @Silverstein:2015er], indicating that bleaching may have adaptive value in allowing corals to rapidly acclimatize to changing conditions. Moreover, this phenomenon could be leveraged as a way to increase the resilience of reef corals through directed manipulation of *Symbiodinium* communities [@Baker:2008p3421].

*Symbiodinium* is a highly diverse genus comprised of nine clades [@Pochon:2014bk], which are further subdivided into numerous types [with species descriptions underway; @LaJeunesse:2014io; @Wham:2017dy] with varying functional biology [@Suggett:2015dv]. *S. trenchii*, a member of *Symbiodinium* clade D, is a heat-tolerant symbiont that confers a bleaching threshold ~1-2°C higher than other *Symbiodinium* types [@Silverstein:2017ep]. Increases in the abundance of clade D symbionts have been observed during and after bleaching in several coral species [@Jones:2008p113; @LaJeunesse:2009p7364], and increases in clade A have been observed in others [@Grottoli:2014jo]. Moreover, the vast majority of coral species surveyed have been observed in association with *Symbiodinium* from multiple clades [@Silverstein:2012ub], suggesting the potential for symbiont shuffling is widespread among corals.

Nevertheless, the mechanisms governing the process of symbiont shuffling, and why it readily occurs in some corals but not others, are still poorly understood. In *Orbicella faveolata*, symbiont shuffling was shown to depend on the severity of bleaching (i.e., the magnitude of disturbance to the existing community), as well as the environmental conditions during recovery [@Cunning:2015ja], suggesting that the relative performance of different symbionts influences the trajectory of symbiont community re-establishment. However, general rules that link symbiont performance to variability in symbiont shuffling among coral species with different *Symbiodinium* types have never been identified. Ultimately, a deeper understanding of symbiont community dynamics in corals is a critical foundation for 1) predicting corals responses to environmental change [@Logan:2013hp], and 2) manipulating symbiont communities as a way to increase resilience in coral reef restoration [@vanOppen:2015es].

Here, we investigated symbiont shuffling through controlled bleaching experiments with three Caribbean coral species hosting mixed *Symbiodinium* assemblages: *O. faveolata* (with clades B and D), *S. siderea* (with clades C and D), and *M. cavernosa* (with clades C and D). We used chlorophyll fluorescence in combination with clade-specific quantitative PCR assays to explore links between the photochemical efficiencies of co-occurring symbionts and the magnitude of symbiont shuffling observed after bleaching.

# Materials and Methods

## Experimental manipulations

Two separate bleaching and recovery experiments were conducted. The first experiment included both *O. faveolata* (`r nOfavcol` colonies, `r min(rangeOfav)`-`r max(rangeOfav)` cores colony^-1^, n=`r nOfavcore` cores) and *S. siderea* (`r nSsidcol` colonies, `r min(rangeSsid)`-`r max(rangeSsid)` cores colony^-1^, n=`r nSsidcore` cores), with bleaching treaments of exposure to 32°C for 7, 10, and 14 days [@Cunning:2015ja], and recovery for 3 months at 24 or 29°C. The second experiment was conducted on *M. cavernosa* (`r nMcavcol` colonies, `r min(rangeMcav)`-`r max(rangeMcav)` cores colony^-1^, n=`r nMcavcore` cores), with a bleaching treatment of 32°C for 10 days [@Silverstein:2015er] and recovery for 3 months at 24 or 29°C. Prior to experimentation, the *O. faveolata* cores contained varying mixtures of *Symbiodinium* in clades B and D [@Cunning:2015ja], while *S. siderea* colonies contained mixtures of clades C and D [@Cunning:2013tl]. *M. cavernosa* contained only clade C *Symbiodinium* upon collection, but an initial bleaching and recovery experiment involving some of these cores produced a range of mixtures of clades C and D [@Silverstein:2015er]. Both experiments were conducted in the same indoor, semi-recirculating coral culture facility at the University of Miami. For additional details on coral collection and experimental set-up, see @Silverstein:2015er and @Cunning:2015ja.

## Sampling and data collection

In both experiments, maximum photochemical efficiency (*F~v~/F~m~*) of *Symbiodinium* within each coral core was measured prior to heat stress and at the end of heat stress using an Imaging PAM (Walz, Effeltrich, Germany). At the same time points, small tissue biopsies were taken from each core for genomic DNA extraction and quantification of the relative abundance of *Symbiodinium* clades using qPCR. Tissue samples were collected again after 3 months of recovery from bleaching to analyze changes in *Symbiodinium* community structure (i.e., symbiont shuffling). For additional details on sampling and data collection, see @Silverstein:2015er and @Cunning:2015ja.

## Data analysis

Photochemical efficiency (*F~v~/F~m~*) was analyzed by a linear mixed model (*lmer*; [@Bates:2015wh]) with fixed effects of time (pre- or post-heat-stress), proportion D1a (measured at the same time point), and coral species, and with random intercepts for each colony. Backward elimination of non-significant effects from a fully crossed model was performed using partial F-tests in *lmerTest* [@lmerTestTestsinL:j0go4ZGL]. By relating the photochemical efficiency of *Symbiodinium* communities to their taxonomic composition [*sensu*  @Suggett:2009p4201], we can tease apart the relative photochemical performance of the two different symbionts in each coral. Specifically, regression coefficients for the proportion D1a were interpreted as the photochemical advantage of D1a for each species under ambient (pre-heat-stress; *A^D^~Amb~*) and heated (post-heat-stress; *A^D^~Heat~*) conditions, and tested for statistical differences by pairwise comparisons using *lsmeans* [@Lenth:2015vl]. Stress duration (i.e., 7, 10, or 14 days) was not included as a predictor since it had no impact on the photochemical advantage of D1a in the first experiment (i.e., no significant interaction between stress duration and proportion D1a).

Symbiont shuffling was analyzed by modeling the final proportion of D1a (after 3 months of recovery) as a function of the initial proportion D1a (before bleaching), using a quasi-binomial generalized linear model (GLM) with species, stress duration, and recovery temperature as co-predictors. Fitted values for the final proportion of D1a for each species at each stress duration (averaged across recovery temperatures, and predicted for *M. cavernosa* at 7- and 14-day durations using the *effects* package [@Fox:2003uu]), were plotted against initial proportion of D1a with 90% confidence intervals. From these fitted values, an integrated metric of symbiont shuffling was generated across the full range of initial D1a proportions by finding the area between the fitted curve for each group and the diagonal line of identity, which indicates no change in symbiont proportions. This metric was scaled to a range of -1 (e.g., complete loss of D1a) to +1 (e.g., complete dominance by D1a), with zero indicating no symbiont shuffling.

Finally, we used a quasi-binomial GLM to model the integrated symbiont shuffling response as a function of the photochemical advantage of D1a under heat stress (A^D^~Heat~) for each species, and days of exposure to heat stress. All data analysis was conducted using R v3.3.1 [@RALanguageandEn:2014wf]. Data and R code to reproduce all analyses and figures presented here are available at: github.com/jrcunning/symshuff-3c.

# Results

## Photochemistry of symbiont assemblages

Measured values of *F~v~/F~m~* were significantly affected by the proportion of D1a in symbiont assemblages (Fig. 1), but this effect varied among coral species and between ambient and heat-stress conditions (three-way interaction, *F*=`r round(lmerTest::anova(bleachmod)$F[7], 3)`, *p*=`r round(lmerTest::anova(bleachmod)$P[7], 3)`). In all three coral species, *F~v~/F~m~* was negatively related to the proportion of D1a under ambient conditions, but positively related to the proportion of D1a under heat stress. The slope of each line in Fig. 1 (i.e., the difference in *F~v~/F~m~* when proportion D=1 and when proportion D=0) can be interpreted as the photochemical advantage of D1a compared to the other symbiont (Fig. 2). The advantage of D under ambient conditions (*A^D^~Amb~*) was significantly more negative (i.e., a greater disadvantage) in *O. faveolata* (`r round(AD[4,3], 3)`) compared to *M. cavernosa* (`r round(AD[6,3], 3)`, *p*=`r round(lst.pw$p.value[5], 3)`), and intermediate (`r round(AD[5,3], 3)`, but not significantly different, *p*>`r round(min(lst.pw$p.value[c(4,6)]), digits=1)`) in *S. siderea*. Under heat stress, the advantage of D1a (*A^D^~Heat~*) was greatest in *M. cavernosa* (`r round(AD[3,3], 3)`), intermediate in *S. siderea* (`r round(AD[2,3], 3)`), and lowest in *O. faveolata* (`r round(AD[1,3], 3)`, all pairwise comparisons, *p*<`r plyr::round_any(max(lst.pw$p.value[c(1,2,3)]), 0.001, ceiling)`).

![Figure 1. Photochemical efficiency as a function of the proportion of D1a *Symbiodinium* in **A)** *O. faveolata*, **B)** *S. siderea*, and **C)** *M. cavernosa*. Solid lines represent ambient conditions, and dotted lines represent the relationship under heat stress (32$^\circ$C). Shaded regions represent 90% confidence intervals around fitted values.](../figures/Fig1.png)

![Figure 2. Photochemical advantage of D1a symbionts in each coral species under ambient conditions (darker bars) and under heat stress (lighter bars). Values indicate the difference in Fv/Fm of corals with all and no D1a symbionts (i.e., slopes of the lines in Fig. 1), with error bars indicating 95% confidence intervals.](../figures/Fig2.png)

## Symbiont shuffling

The proportion of D1a after recovery from bleaching (Fig. 3A) depended on the initial proportion of D1a (*F*=`r round(dropterm(propDmod, test="F")$F[2], 3)`, *p*`r format.pval(dropterm(propDmod, test="F")$P[2], eps=0.0001)`) and an interaction among coral species, heat stress duration, and recovery temperature (*F*=`r round(dropterm(propDmod, test="F")$F[3], 3)`, *p*=`r round(dropterm(propDmod, test="F")$P[3], 3)`). Fitted values were averaged across recovery temperatures in order to focus on differences among species and heat stress durations. Across the full range of initial D1a proportions, the integrated symbiont shuffling response, $S$ (Fig. 3B), showed a loss of D1a in *O. faveolata* after recovery from 7 and 10 days of heat stress ($S$=`r round(auc[4,3], 3)` and `r round(auc[7,3], 3)`), but a gain of D1a after recovery from 14 days of heat stress ($S$=`r round(auc[1,3], 3)`). In contrast, *S. siderea* underwent greater shuffling toward D1a after 7 ($S$=`r round(auc[5,3], 3)`), 10 ($S$=`r round(auc[8,3], 3)`), and 14 ($S$=`r round(auc[2,3], 3)`) days of heat stress. *M. cavernosa* shuffled almost completely to D1a assemblages following 10 days of stress ($S$=`r round(auc[9,3], 3)`), and thus the statistical model also predicted near complete shuffling to D after 7 ($S$=`r round(auc[6,3], 3)`) and 14 ($S$=`r round(auc[3,3], 3)`) days of heat stress.

![Figure 3. Symbiont shuffling in three coral species. Panel **(A)** shows the proportion of D1a *Symbiodinium* before bleaching and after recovery, with the diagonal line of identity representing no change in symbiont proportions. Fitted values for each species (indicated by color) and heat stress duration (indicated by line type) are plotted with shaded areas representing 95% confidence intervals. Panel **(B)** shows the integrated symbiont shuffling response (from before bleaching to after recovery) for each species and bleaching treatment, calculated as the area above or below the diagonal line of identity, normalized to a range of -1 (i.e., a complete loss of D1a symbionts) to +1 (i.e., complete dominance by D1a symbionts).](../figures/Fig3.png)

## Relationship between photochemistry and symbiont shuffling

The magnitude of symbiont shuffling showed a significant statistical relationship (Fig. 4A) with both the photochemical advantage of D1a under heat stress (*A^D^~Heat~*, *F*=`r round(anova(glmod, test="F")$F[2], 1)`, *p*`r format.pval(anova(glmod, test="F")$P[2], eps=0.0001)`) and the duration of heat stress (*F*=`r round(anova(glmod, test="F")$F[3], 1)`, *p*`r format.pval(anova(glmod, test="F")$P[3], eps=0.0001)`). Specifically, increases in the amount of D1a were greater for higher values of *A^D^~Heat~* and longer durations of heat stress. The model predicts the duration of heat stress required to cause symbiont shuffling toward D1a for different values of *A^D^~Heat~* (Fig. 4A), and shows that 7 days of heat stress results in shuffling toward D1a for *A^D^~Heat~*>~0.06. Lower values of *A^D^~Heat~* cause a decrease in the proportion of D1a after recovery, even following 14 days of heat stress for *A^D^~Heat~*<~0.01.

Across these three coral species, the interaction of *A^D^~Heat~* and the duration of heat stress explains `r round(pseudoR2*100, 1)`% of the variability in the symbiont shuffling response (Fig. 4B).

![Figure 4. Symbiont shuffling as a function of the photochemical advantage of D1a and heat stress duration. Panel **(A)** shows the symbiont shuffling response (as in Fig. 3B) as a function of heat stress duration (days) and the photochemical advantage of D1a under heat stress (A^D^~Heat~), with lines corresponding to the A^D^~Heat~ values for the three coral species studied (indicated by colored lines), and other (unobserved) values (gray lines). Panel **(B)** shows that the symbiont shuffling responses observed across all three species are predicted well (pseudo-R^2^=0.93) by the time-integrated photochemical advantage of D1a under heat stress.](../figures/Fig4.png)

# Discussion

These experiments revealed significant differences in the photochemical efficiencies of co-occurring symbiont clades in mixed symbiont communities. At ambient temperatures, D1a symbionts had significantly lower *F~v~/F~m~* than the other symbiont clade within each coral host. Since ambient temperatures are not expected to cause photodamage, the reduced photochemical efficiency of D1a may be due to reduced nutrient sufficiency [@Parkhill:2001eu], greater pigment antenna size or association with PSI [@Suggett:2009p4201], greater excitation 'spillover' from PSII to PSI [@Slavov:2016kh], or other mechanisms of non-photochemical quenching [@Osmond:1994ua]. Indeed, inherent differences in photobiological traits produce taxonomic variation in *F~v~/F~m~* among broad microalgal classes [@Suggett:2009p4201] as well as different types of *Symbiodinium* [@Suggett:2008p960, @Suggett:2015dv]. Regardless of the underlying mechanisms, the reduced photochemical efficiency of D1a relative to other *Symbiodinium* *in hospite* suggests D1a has a reduced potential for PSII photochemistry. This disadvantage may explain why D-dominated corals tend to grow more slowly [@Little:2004p140], and why D1a may be displaced by other symbionts in the absence of stress [@Thornhill:2006p3638; @LaJeunesse:2009p7364].

However, under heat stress, the relative performance of symbionts was reversed, with D1a maintaining a higher photochemical efficiency than each co-occurring symbiont. These reductions in *F~v~/F~m~* can be attributed to accumulation of damage impacting photosystem II, which is known to vary among *Symbiodinium* types [@Warner:1996p8115]. Indeed, clade D *Symbiodinium* have been shown to maintain higher *F~v~/F~m~* under heat stress [@Rowan:2004p3383, @Silverstein:2015er], indicative of greater thermal tolerance. However, the magnitude of D1a's advantage under heat stress differed among the three coral species, being greatest in *M. cavernosa*, intermediate in *S. siderea*, and lowest in *O. faveolata*.

The differential performance of different symbionts in different coral hosts may be explained by variation in symbiont-specific photobiological traits, and/or host-specific traits that modulate symbiont performance. For example, coral tissue architecture may influence internal light and nutrient microenvironments [@Wangpraseurt:2015hh], and host regulation of nitrogen and/or carbon dioxide supply to symbionts may influence their performance [@Tansik:2017ct]. In both *O. faveolata* and *S. siderea*, D1a symbionts showed the same *F~v~/F~m~*, indicating that the different relative advantages of D1a were in fact due to differential performance of the co-occurring symbionts. D1a in *M. cavernosa* had slightly lower *F~v~/F~m~* under ambient conditions (and higher under heat stress) compared to D1a in the other two corals. However, since this followed repeated bleaching in a separate experiment for *M. cavernosa*, we are unable to attribute these differences to unique host traits, the influence of prior bleaching and recovery, or random experimental effects.

Nevertheless, we could test whether the relative photophysiological performance of the two symbionts within each coral host was predictive of the magnitude of symbiont shuffling within that host following heat stress. Indeed, 93% of the variability in symbiont shuffling was predicted by *A^D^~Heat~* and the duration of heat stress applied. Specifically, greater increases in D1a occurred when *A^D^~Heat~* was higher, and when heat stress was longer. Symbiont assemblages in which D had less of an advantage under stress (e.g. *O. faveolata*) required more stress to induce shuffling to D1a, while assemblages in which D1a had a large advantage (e.g. *M. cavernosa*) required less stress. This finding suggests that both the relative sensitivity to stress and the duration of stress determine symbiont shuffling, similar to the way that both concentration and duration of exposure influence a pharmacological response [@Miller:2000df].

This statistical result reveals predictors of symbiont shuffling but not necessarily the mechanisms underlying them. However, we speculate that if D1a performs better under at 32°C, it may be expelled at lower rates than co-occurring symbionts [@Silverstein:2017ep], or may even continue to proliferate. Furthermore, after heat stress is alleviated (but corals have bleached), amplified internal light environments characteristic of bleached corals [@Wangpraseurt:2012cu] may sustain the photochemical advantage of D1a, allowing it to proliferate at higher rates as the symbiont population regrows. Only once symbionts reach high densities may the internal microenvironment (e.g., darker, more nutrient-scarce conditions) again put D1a at a disadvantage, explaining why D may be gradually replaced over time in the absence of stress [@Thornhill:2006p3638]. Importantly, the rate of compositional change in the scenario outlined here would be proportional to the relative difference in the advantage of one symbiont over another, consistent with our observations across coral species.

Whether these same rules may apply in other *Symbiodinium* communities, particularly those without clade D, but which demonstrate shuffling between other types [e.g., *Porites* with clades A and C; @Grottoli:2014jo], remains to be investigated. Other populations of *O. faveolata* may shuffle variously among clades A, B, C, and D [@Baker:1999p147; Jones et al., Serrano et al., unpublished). In contrast, symbiont shuffling appears to be rare in other coral taxa, such as *Pocillopora* in the Mexican Pacific [@McGinley:2012jn] and *Montipora capitata* in Hawaii [@Cunning:2016jb], despite the fact that these corals host members of *Symbiodinium* clades C and D that differ greatly in their sensitivity to stress. While the links identified here may not explain symbiont community dynamics in all cases, this study expands our understanding of symbiont shuffling in a variety of important Caribbean reef-building species. A better understanding of this phenomenon will allow us to refine predictive models of coral responses to environmental change [@Logan:2013hp], as well as methods to manipulate symbiont communities in corals, which may be an increasingly important tool in coral reef conservation [@vanOppen:2015es; @Aswani:2015cq].

# Acknowledgements
RC and RNS were supported by NSF Graduate Research Fellowships and University of Miami Graduate Fellowships, with additional funding from RSMAS student awards, the Captain Harry D. Vernon Memorial Scholarship, and the Garden Club of America. Additional support was provided by NSF (OCE-0547169 and OCE-1358699), the Lenfest Ocean Program, a Pew Fellowship in Marine Conservation, and a University of Miami Provost Research Award in Natural Science and Engineering to ACB. We thank Z Schwartz, K Ondrasik, K Dziedzic, N Formel, K Montenero, P Jones, R Winter, R Okazaki, L Gordon, and N. Guy for laboratory assistance.

# References